### 2021-05-26 Paper.tex
modified

\begin{table}[H]
\caption{List of benchmarks implemented for each feature of the code.}
\centering
\begin{tabular}{| l | l | l |}
\toprule
 \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
\midrule
  Solver  & Stokes flow & \ref{sec:stokes} \\
\hline
  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk} & \multirow{2}{*}{\ref{sec:runge}} \\
  markers advection  &  & \\
\hline
  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
    & Falling block & \ref{sec:block} \\
\hline
    & 2D Stokes sphere & \ref{sec:sphere} \\
    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
    free surface & Topography relaxation & \ref{sec:crameri} \\
  & Spontaneous subduction & \ref{sec:subduction} \\
\hline
  Erosion and sedimentation  & - & - \\
\hline
   Non-linear  & Slab detachment & \ref{sec:slab} \\
   visco-plasticity & Indenter & \ref{sec:indenter} \\
    & Brick & \ref{sec:brick} \\
\hline
  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
    & Simple shear heating & \ref{sec:simple_shear} \\
    & Shear and adiabatic heating & \ref{sec:shear} \\
\hline
\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
& Thin layer entrainment & \ref{sec:thin} \\
\hline
  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
  and hydration  &  & \\
\hline
  Melting & Experimental melting curves & \ref{sec:katz} \\
\bottomrule
\end{tabular}
\label{tab:list}
\end{table}

@@

Other important quantities calculated in some benchmarks are the root-mean-square velocity and the Nusselt number, calculated as
\[v_{\textrm{rms}}=\sqrt{\frac{1}{V_{\Omega}}\int_{\Omega}|\bm{v}|^2 d\Omega}\]
and
\[\textrm{Nu}=\frac{\int_{top}\bm{\nabla} T \cdot \bm{n} ds}{\int_{bottom}T ds}\]
respectively.

@@

The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
Fig. \ref{fig:MUMPS}.

@@

\subsection{Markers advection}\label{sec:runge}
The 4th-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
in a unit square domain with different grid resolution (\num{100x100} and \num{200x200} elements) and 1000000 markers. The Courant number varies between 0.1
and 0.5 and the velocity field is prescribed in the entire domain as
\begin{eqnarray}
u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
\end{eqnarray}
At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge}).
As we expect, markers are back to their initial location after a $2\pi$ rotation ($t=1$, red, green and blue in Fig. \ref{fig:runge}), with an error of the
final position that increase moving away from the centre, more evident in case of higher time steps or lower grid resolution (Fig. \ref{fig:runge_err}).

@@

The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
correlation with the pressure field.

@@

\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
We distinguish three different types of velocity boundary conditions:
\begin{enumerate}
\item FS: free slip conditions on all sides.
\item NS: no slip conditions on all sides.
\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
\end{enumerate}
All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}. Results in terms of the $v_{\textrm{rms}}$ for NS model are also shown
in Fig. \ref{fig:inst_sphere}.

@@

$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps.

@@

The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}).

@@

\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}. Results in terms of the $v_{\textrm{rms}}$ in case of an arithmetic average are also
shown in Fig. \ref{fig:inst_spherefs}.

@@

The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}).

@@

The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}).

@@

The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}.

@@

The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}.

@@

Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}.

@@

As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}).

@@

Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}).

@@

\subsection{Shear and adiabatic heating}\label{sec:simple_shear}
This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
\[T(t)=H_s t\]
In this case we have
\begin{eqnarray}
\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
\end{eqnarray}
and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]

The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}).

@@

As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}).

@@

The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
with lower errors for higher resolution grids (Table \ref{tab:mantle}).

@@

For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}.

@@

\begin{table}[H]

@@

As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively).

>>>>

\begin{table}[H]
\caption{List of benchmarks implemented for each feature of the code. Data of all benchmarks can be found at 
\url{https://github.com/aleregorda/Benchmarks}}
\centering
\begin{tabular}{| l | l | l |}
\toprule
 \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
\midrule
  Solver  & Stokes flow & \ref{sec:stokes} \\
\hline
  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk} & \multirow{2}{*}{\ref{sec:runge}} \\
  markers advection  &  & \\
\hline
  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
    & Falling block & \ref{sec:block} \\
\hline
    & 2D Stokes sphere & \ref{sec:sphere} \\
    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
    free surface & Topography relaxation & \ref{sec:crameri} \\
  & Spontaneous subduction & \ref{sec:subduction} \\
\hline
  Erosion and sedimentation  & - & - \\
\hline
   Non-linear  & Slab detachment & \ref{sec:slab} \\
   visco-plasticity & Indenter & \ref{sec:indenter} \\
    & Brick & \ref{sec:brick} \\
\hline
  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
    & Simple shear heating & \ref{sec:simple_shear} \\
    & Shear and adiabatic heating & \ref{sec:shear} \\
\hline
\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
& Thin layer entrainment & \ref{sec:thin} \\
\hline
  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
  and hydration  &  & \\
\hline
  Melting & Experimental melting curves & \ref{sec:katz} \\
\bottomrule
\end{tabular}
\label{tab:list}
\end{table}

@@

Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
\begin{eqnarray}
v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
\end{eqnarray}
respectively, top and bottom Nusselt numbers
\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
the average rate of work done against gravity
\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
the average rate of viscous dissipation
\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]

@@

The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.

@@

\subsection{Markers advection}\label{sec:runge}
The 4th-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
in a unit square domain with different grid resolution (\num{100x100} and \num{200x200} elements) and 1000000 markers. The Courant number varies between 0.1
and 0.5 and the velocity field is prescribed in the entire domain as
\begin{eqnarray}
u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
\end{eqnarray}
At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge}).
As we expect, markers are back to their initial location after a $2\pi$ rotation ($t=1$, red, green and blue in Fig. \ref{fig:runge}), with an error of the
final position that increase moving away from the centre, more evident in case of higher time steps or lower grid resolution (Fig. \ref{fig:runge_err}). 
All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Zalesak_Disk}.

@@

The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.

@@

\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
We distinguish three different types of velocity boundary conditions:
\begin{enumerate}
\item FS: free slip conditions on all sides.
\item NS: no slip conditions on all sides.
\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
\end{enumerate}
All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
are also shown in Fig. \ref{fig:inst_sphere}.

@@

$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.

@@

The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.

@@

\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.

@@

The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.

@@

The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.

@@

The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.

@@

The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.

@@

Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.

@@

As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.

@@

Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.

@@

\subsection{Shear heating}\label{sec:simple_shear}
This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
\[T(t)=H_s t\]
In this case we have
\begin{eqnarray}
\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
\end{eqnarray}
and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]

The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.

@@

As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.

@@

The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.

@@

For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.

@@

\begin{table}

@@

As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
